Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-07-01 Manuscript/manuscript.Rmd
edits

![**The proportional importance of optimally built cell line specific models was outweighed by transcription factor binding sites across resolutions.** Stacked barplots illustrate the proportion of total predictive importance in cell line specific random forest classifiers for (A) GM12878 and (B) K562 respectively. Model inputs included the optimal data level characteristics determined by the ensemble framework (SMOTE re-sampling; with distance-type predictors). Predictive importance was averaged for models built on chromosome specific regions of the linear genome. Proportional importance was then calculated according to the annotation group of the feature space including chromatin state segmentations (CSS), histone modifications (HM), and transcription factor binding sites (TFBS).](Figures/proportion_of_importances_v2.png)

![**Optimally built TFBS-specific models outperformed other annotation-specific models across resolution for GM12878** Performances were compared for random forest models built using specific groups of annotations as the feature space including: histone modifications (HM; red), chromatin state segmentations (CSS; green), transcription factor binding sites (TFBS; blue), and a feature space containing all 3 types combined (combined; purple). Each bar represents a model ran through the ensemble framework, with re-sampling technique and predictor-type characteristics fixed to SMOTE and distance-types. The error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosome.](Figures/figure6_comparing_sets_of_annotations_gm12878.png)

@@

![**Top 20 most predictive transcription factor binding sites sorted by 10 kb for GM12878.** A heatmap illustrating the importance of the key cell line specific transcription factors for optimally built random forest models across resolution. Importances were sorted according to the 10 kb resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red.](Figures/figure7_variable_importance_heatmaps_tfbs_gm12878.png)

>>>>

<!--
![**The proportional importance of optimally built cell line specific models was outweighed by transcription factor binding sites across resolutions.** Stacked barplots illustrate the proportion of total predictive importance in cell line specific random forest classifiers for (A) GM12878 and (B) K562 respectively. Model inputs included the optimal data level characteristics determined by the ensemble framework (SMOTE re-sampling; with distance-type predictors). Predictive importance was averaged for models built on chromosome specific regions of the linear genome. Proportional importance was then calculated according to the annotation group of the feature space including chromatin state segmentations (CSS), histone modifications (HM), and transcription factor binding sites (TFBS).](Figures/proportion_of_importances_v2.png)
-->

<!--
![**Optimally built TFBS-specific models outperformed other annotation-specific models across resolution for GM12878** Performances were compared for random forest models built using specific groups of annotations as the feature space including: histone modifications (HM; red), chromatin state segmentations (CSS; green), transcription factor binding sites (TFBS; blue), and a feature space containing all 3 types combined (combined; purple). Each bar represents a model ran through the ensemble framework, with re-sampling technique and predictor-type characteristics fixed to SMOTE and distance-types. The error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosome.](Figures/comparing_sets_of_annotations_v2.png)
-->

@@

<!--
![**Top 20 most predictive transcription factor binding sites sorted by 10 kb for GM12878.** A heatmap illustrating the importance of the key cell line specific transcription factors for optimally built random forest models across resolution. Importances were sorted according to the 10 kb resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red.](Figures/figure7_variable_importance_heatmaps_tfbs_gm12878.png)
-->

![**Assessing optimally built cell line specific random forest classifiers for predicting TAD boundary formation.** (A) Comparing the proportion of total predictive importance carried by each group of functional genomic annotation across resolution. Predictive importance was averaged for models built on chromosome specific regions of the linear genome. Proportional importance was then calculated according to the annotation group of the feature space including chromatin state segmentations (CSS), histone modifications (HM), and transcription factor binding sites (TFBS). (B) Comparing models using specific groups of functional genomic annotation as the feature space in addition to a model with a feature space containing all 3 types combined (combined; purple). Error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosomal section of the linear genome. (C) Heatmaps illustrating the importance of the top 20 cell line specific transcription factors across resolution. Importances were sorted according to the row-wise average across resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red and dark blue for the GM12878 and K562 cell line respectively.](Figures/prop_imps_a_comparing_annots_b_var_imps_c.png)
